---
title: "words_rm function"
author: "Simone D'Ambrogio"
date: "24/1/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## words_rm
If a variable is composed by numbers and letters, it will be converted by R in factor (or character). However, sometimes this make us deal with meaningless variables. 

We ran into a case like this yesterday, whith the "Age" variable that conteined numbers and ... words.
The first and more psychological question is: what's the problem with these guys?? A second question, this time more technical and maybe more boring is: how can we solve this problem efficiently and in a way that minimize the probability of error? In my opinion, put the hands directly on the excel file doesn't satisfy neither of the two criteria.

Trying to solve this problem, yesterday I created a small function that hopefully may come in handy.

```{r cars}
## Open the dataset

#setwd(...)
sa <- read.csv(file = "JDMLab_Scarcity Ambientale .csv")
str(sa$Age)
#As we can see, the Age variable is a factor and cannot be converted to numbers as long as 
#there are letters
```

```{r}
#Convert the Age variable in character
sa$Age <- as.character(sa$Age)

library(tidyverse)
#Detect the values that in the Age variable that contain letters. 
#Our goal is to keep the in numbers and discard the letters
unique(unlist(sapply(c(letters,LETTERS), function(i) sa$Age[str_detect(sa$Age, i)])))
```

## words_rm function

```{r}
#This is the function
words_rm <- function(data, variable, words_to_rm, replace = ""){
  
  library(tidyverse)
  
  if(!require(tidyverse)){
    install.packages("tidyverse")
    library(tidyverse)
  }
  
  if(!is.character(data[, variable])) stop("The arg variable must be a character!")
  
    new_number <- sapply(1:length(data[, variable][str_detect(data[, variable], words_to_rm)]), 
                       function(i){
                         data[, variable][str_detect(data[, variable], words_to_rm)][i] <- 
                           gsub(words_to_rm, replace, data[, variable][str_detect(data[, variable], words_to_rm)][i])
                       })
  
  data[str_detect(data[, variable], words_to_rm),variable] <- new_number
  return(data[, variable])
}
```

See how the function works in pracice

```{r pressure, echo=FALSE}
#The words_rm function eliminates the worlds that we want to eliminate
#There are only 4 arguments:
#data: name of the dataset
#variable: character variable that we want to modify. don't forget the quotation marks
#words_to_rm: words that we want to eliminate separated by |
#replace: numbers or words we want instead of words_rm. Default is "" and you should 
#keep it if you want to eliminate the words

sa$Age <- words_rm(data=sa, variable="Age", words_to_rm = "anni|anno")

#Detect again the values that in the Age variable that contain letters. 
unique(unlist(sapply(c(letters,LETTERS), function(i) sa$Age[str_detect(sa$Age, i)])))

#The values containing words are now redused from 12 to 3. 
```



```{r}
#If we want to convert the one who answered Sessantaqattro instead of 64, we
#can use the argument replace as below
sa$Age <- words_rm(data=sa, variable='Age', words_to_rm = "Sessantaquattro", replace = "64")

#Detect again the values that in the Age variable that contain letters. 
unique(unlist(sapply(c(letters,LETTERS), function(i) sa$Age[str_detect(sa$Age, i)])))
```

I hope you find it useful. If you find bugs or if you have question or curiosities, please contact me :)

Simone